---
title: "Creating Balanced Panels from the GWL Dataset"
output: pdf_document
---

First course of action is to load our data, saved in the feather format.
```{r}
# install.packages(c('dplyr', 'feather', 'ggplot2', 'rmarkdown', 'lubridate'))
library(dplyr)
library(feather)
library(ggplot2)
library(rmarkdown)
library(lubridate)

gwl_2023_24 <- read_feather('data/gwl_23_24_filtered.feather')
glimpse(gwl_2023_24)
```

Second course of action will be to find how many stations made ≤ median count of observations over the year.
```{r}
obs_per_station <- gwl_2023_24 %>%
  group_by(latitude, longitude) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

obs_per_station_median <- obs_per_station %>% filter(count <= 3)
glimpse(obs_per_station_median)
```
11208/18849 (~60%) stations made ≤median(3) observations.
We will now filter the observations from those stations.
```{r}
gwl_median <- gwl_2023_24 %>%
  left_join(obs_per_station, by = c('latitude', 'longitude')) %>%
  filter(count <= 3)

glimpse(gwl_median)
```
`gwl_median` stores the observations from stations who made ≤3 observations in the period May 2023-May 2024.
```{r}
# observations per station among stations which made ≤median count of observations.
obs_per_station_median <- gwl_median %>%
  group_by(latitude, longitude) %>%
  summarise(n_observations = n())

# silly naming to find out the number of stations which made certain observation counts
freq_freq <- obs_per_station_median %>%
  group_by(n_observations) %>%
  summarise(stations_with_n_observations = n())

glimpse(freq_freq)
```
3870 stations made 2 observations across the year.
We will now try to see if the observations from those stations will allow us to create a balanced panel.
```{r}
gwl_2_obs <- gwl_median %>%
  filter(count == 2)

glimpse(gwl_2_obs)
```
The next step to finding a balanced panel would be to first see the monthly frequency of observations
over the 2-observation dataset.
```{r}
monthly_freq <- gwl_2_obs %>%
  mutate(date = as_datetime(data_time),  # Convert POSIX timestamp to datetime
         month = floor_date(date, "month")) %>%
  group_by(month) %>%
  summarise(obs_per_month = n())

glimpse(monthly_freq)

monthly_freq$month <- as.Date(monthly_freq$month)
ggplot(monthly_freq, aes(x = month, y = obs_per_month)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Number of Observations per Month (2-observation Stations)",
       x = "Month",
       y = "Number of Observations") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  geom_label(data = monthly_freq, aes(label = obs_per_month))

ggsave('../../plots/monthly_freq_2_obs.png')
```
The above plot shows that there is potential in creating a balanced panel of observations from stations
where observations were made in the following periods:
- May 2023 and August 2023
- May 2023 and Jan 2024
- May 2023 and Nov 2023
- May 2023 and Dec 2023

We will now proceed to find the number of stations that made observations according to the suggested
panels.
```{r}
gwl_2_obs_may <- gwl_2_obs %>%
  filter(month(data_time) == 5 & year(data_time) == 2023)

gwl_2_obs_aug <- gwl_2_obs %>%
  filter(month(data_time) == 8 & year(data_time) == 2023)

may_aug_locations <- gwl_2_obs_may %>%
  select(latitude, longitude) %>%
  inner_join(gwl_2_obs_aug %>% select(latitude, longitude),
             by = c("latitude", "longitude")) %>%
  distinct()

glimpse(may_aug_locations)
```
There are 2149 stations that made their 2 observations in May 2023 and then in August 2023.
We will employ similar techniques for the other panels.

```{r}
gwl_2_obs_nov <- gwl_2_obs %>%
  filter(month(data_time) == 11 & year(data_time) == 2023)

gwl_2_obs_jan <- gwl_2_obs %>%
  filter(month(data_time) == 1 & year(data_time) == 2024)

gwl_2_obs_dec <- gwl_2_obs %>%
  filter(month(data_time) == 12 & year(data_time) == 2023)

may_jan_locations <- gwl_2_obs_may %>%
  select(latitude, longitude) %>%
  inner_join(gwl_2_obs_jan %>% select(latitude, longitude),
             by = c("latitude", "longitude")) %>%
  distinct()

glimpse(may_jan_locations)
```

```{r}
may_nov_locations <- gwl_2_obs_may %>%
  select(latitude, longitude) %>%
  inner_join(gwl_2_obs_nov %>% select(latitude, longitude),
             by = c("latitude", "longitude")) %>%
  distinct()

glimpse(may_nov_locations)
```
```{r}
may_dec_locations <- gwl_2_obs_may %>%
  select(latitude, longitude) %>%
  inner_join(gwl_2_obs_dec %>% select(latitude, longitude),
             by = c("latitude", "longitude")) %>%
  distinct()

glimpse(may_dec_locations)
```
The number of stations that create a balanced panel for the suggested panels are as follows:
- May 2023 and August 2023 = 2149
- May 2023 and Jan 2024 = 502
- May 2023 and Nov 2023 = 82
- May 2023 and Dec 2023 = 20

We will now create the following datasets out of the groundwater level data for further analysis:
- **Pre-Monsoon to Monsoon** (May 2023 and August 2023)
- **Pre-Monsoon to Post-Monsoon** (May 2023 and Nov 2023/Dec 2023/Jan 2024)

```{r}
# premonsoon_monsoon should be 2149*2 = 4298 rows beyond all doubt.
may_aug <- bind_rows(gwl_2_obs_may, gwl_2_obs_aug)
glimpse(may_aug_locations)

premonsoon_monsoon <- may_aug %>%
  mutate(data_time = floor_date(data_time, unit = 'month')) %>%
  inner_join(may_aug_locations, by = c('latitude', 'longitude'))


glimpse(premonsoon_monsoon)
write_feather(premonsoon_monsoon, '../../data/premonsoon_monsoon.feather')
```
Correct dimensions confirmed.
Moving onto the pre-monsoon-post-monsoon dataset.
```{r}
# premonsoon_postmonsoon should be (502 + 82 + 20)*2 = 1208 rows beyond all doubt.
d2_locations <- bind_rows(may_nov_locations, may_dec_locations, may_jan_locations)
d2_unclean <- bind_rows(gwl_2_obs_may, gwl_2_obs_nov, gwl_2_obs_dec, gwl_2_obs_jan)

premonsoon_postmonsoon <- d2_unclean %>%
  mutate(data_time = floor_date(data_time, unit = 'month')) %>%
  inner_join(d2_locations, by = c('latitude', 'longitude'))

glimpse(premonsoon_postmonsoon)
write_feather(premonsoon_postmonsoon, '../../data/premonsoon_postmonsoon.feather')
```
Correct dimensions confirmed once again. This will conclude dataset creation on stations with 2 observations
per year.
One more direction worth pursuing is the data associated with stations which were observed _3 times a year_.
The suggested panel is of the structure May 2023 -> Aug 2023 -> (Nov 2023-Jan 2024).
As usual the first course of action would be to see how many stations actually follow this pattern.
```{r}
gwl_3_obs <- gwl_median %>%
        filter(count == 3)

# all observations made in months of interest
gwl_d3_unclean <- gwl_3_obs %>%
  mutate(data_time = floor_date(data_time, unit = 'month')) %>%
  filter(data_time == as.Date('2023-05-01') | data_time == as.Date('2023-08-01') |
           data_time == as.Date('2023-11-01')| data_time == as.Date('2023-12-01')|
           data_time == as.Date('2024-01-01'))

glimpse(gwl_d3_unclean)
monthly_freq <- gwl_d3_unclean %>%
  mutate(date = as_datetime(data_time),  # Convert POSIX timestamp to datetime
         month = floor_date(date, "month")) %>%
  group_by(month) %>%
  summarise(obs_per_month = n())

glimpse(monthly_freq)
```
Promising numbers. How many stations can fit in our panel?
```{r}
t1_stations <- gwl_d3_unclean %>%
  filter(data_time == as.Date('2023-05-01')) %>%
  select(latitude, longitude) %>%
  distinct() %>%
         mutate(sid = paste(latitude, longitude, sep = '-'))

t2_stations <- gwl_d3_unclean %>%
  filter(data_time == as.Date('2023-08-01')) %>%
  select(latitude, longitude) %>%
  distinct() %>% mutate(sid = paste(latitude, longitude, sep = '-'))

t3_stations <- gwl_d3_unclean %>%
        filter(data_time == as.Date('2023-11-01') | data_time == as.Date('2024-01-01')) %>%
        select(latitude, longitude) %>%
        distinct() %>% mutate(sid = paste(latitude, longitude, sep = '-'))

# commons contains all the stations that made observations according to our panel
commons <- inner_join(t1_stations, t2_stations, by = 'sid') %>%
        rename(latitude = "latitude.x", longitude = "longitude.x") %>%
        select(latitude, longitude, sid) %>%
        inner_join(t3_stations, by = 'sid') %>%
        rename(latitude = "latitude.x", longitude = "longitude.x") %>%
        select(latitude, longitude, sid)
glimpse(commons)
```
Let us now create the D3 dataset.
```{r}
# expected length = 964 + 964 + (964) = 2892
gwl_d3_unclean <- gwl_d3_unclean %>%
        mutate(sid = paste(latitude, longitude, sep = '-'))

d3 <- gwl_d3_unclean %>%
        filter(sid %in% commons$sid)
glimpse(d3)
```
Dimensions confirmed.
This will be part of the pre-monsoon-monsoon-post-monsoon analysis (certainly a lot of hyphens.)
```{r}
write_feather(d3, '../../data/full_monsoon.feather')
```