---
title: "Inverse Distance Weighing on Balanced Panels"
output: pdf_document
---

We have acquired D1, D2 and D3 from our work on creating balanced panels. We will now apply some
spatio-temporal analysis techniques.

# Inverse Distance Weighing
We will be using the code provided in _(Wikle et al., 2019)_ to perform inverse distance weighing for
deterministic prediction.
```{r}
# install.packages(c('dplyr', 'fields', 'ggplot2', 'gstat', 'RColorBrewer', 'sp', 'spacetime', 'devtools', 'raster'))
library(dplyr)
library(fields)
library(ggplot2)
library(gstat)
library(RColorBrewer)
library(sp)
library(spacetime)
library(devtools)
library(feather)
library(raster)
library(STRbook)

d1 <- read_feather('../../data/premonsoon_monsoon.feather')
d2 <- read_feather('../../data/premonsoon_postmonsoon.feather')
d3 <- read_feather('../../data/full_monsoon.feather')
dataset <- read_feather('data/gwl_23_24_filtered.feather')
```

This is the function to call IDW.
```{r}
perform_idw <- function(data, time_point, power = 2, resolution = 0.01) {
  data_at_time <- data[data$data_time == time_point, ]

  coordinates(data_at_time) <- ~longitude+latitude

  grid_extent <- extent(
    min(data$longitude) - 0.1,
    max(data$longitude) + 0.1,
    min(data$latitude) - 0.1,
    max(data$latitude) + 0.1
  )

  grid <- raster(grid_extent, resolution = resolution)
  grid_points <- rasterToPoints(grid, spatial = TRUE)

  idw_model <- gstat(
    formula = data_value ~ 1,
    locations = data_at_time,
    nmax = 12,
    set = list(idp = power)
  )
  prediction <- predict(idw_model, grid_points)
  idw_raster <- rasterize(prediction, grid, field = "var1.pred")

  return(list(
    raster = idw_raster,
    prediction = prediction,
    time = time_point
  ))
}

if(!inherits(d1$data_time, "POSIXct")) {
  d1$data_time <- as.POSIXct(d1$data_time)
}

# specific time point for interpolation
time_point <- unique(d1$data_time)[1]

idw_result <- perform_idw(d1, time_point)

raster_df <- as.data.frame(rasterToPoints(idw_result$raster))
colnames(raster_df) <- c("longitude", "latitude", "value")

points_df <- d1[d1$data_time == time_point, c("longitude", "latitude", "data_value")]

ggplot() +
        geom_raster(data = raster_df, aes(x = longitude, y = latitude, fill = value)) +
        geom_point(data = points_df, aes(x = longitude, y = latitude),
                   color = "white", size = 0.5, shape = 21, fill = "white") +
        geom_point(data = arbitrary_points, aes(x = longitude, y = latitude),
                   color = "blue", size = 4, shape = 23, fill = "yellow") +
        scale_fill_viridis_c(name = "Interpolated Value") +
        theme_bw() +  # Use theme_bw() for a non-transparent white background
        theme(
                panel.background = element_rect(fill = "white", color = "black"),  # White panel background
                plot.background = element_rect(fill = "white", color = NA),  # White plot background
                legend.background = element_rect(fill = "white", color = NA)  # White legend background
        ) +
        labs(
                title = paste("Inverse Distance Weighing Interpolation (D1)"),
                subtitle = paste("Time:", format(idw_result$time)),
                x = "Longitude",
                y = "Latitude"
        )

ggsave('../../plots/idw_d1.png', dpi = 300)
```
Let's see.

```{r}
st_idw <- function(dataset, pred_loc, pred_time,
                   coords_cols, time_col, value_col,
                   dist_func = NULL, power = 2,
                   max_dist = Inf, time_scale = 1,
                   cutoff = 3, ...) {

  # If no distance function provided, use default Euclidean distance
  if (is.null(dist_func)) {
    dist_func <- function(x1, y1, t1, x2, y2, t2, time_scale = 1) {
      spatial_dist <- sqrt((x2 - x1)^2 + (y2 - y1)^2)
      temporal_dist <- abs(t2 - t1) * time_scale
      return(sqrt(spatial_dist^2 + temporal_dist^2))
    }
  }

  # Extract coordinates, time and values from dataset
  if (is.character(coords_cols)) {
    x_coords <- dataset[[coords_cols[1]]]
    y_coords <- dataset[[coords_cols[2]]]
  } else {
    x_coords <- dataset[[coords_cols[1]]]
    y_coords <- dataset[[coords_cols[2]]]
  }

  if (is.character(time_col)) {
    times <- dataset[[time_col]]
  } else {
    times <- dataset[[time_col]]
  }

  if (is.character(value_col)) {
    values <- dataset[[value_col]]
  } else {
    values <- dataset[[value_col]]
  }

  # Calculate distances between prediction point and all data points
  distances <- numeric(length(values))
  for (i in seq_along(values)) {
    distances[i] <- dist_func(x_coords[i], y_coords[i], times[i],
                              pred_loc[1], pred_loc[2], pred_time,
                              time_scale = time_scale, ...)
  }

  # Filter points based on max_dist
  valid_indices <- which(distances <= max_dist & !is.na(distances))

  # Check if we have enough points
  if (length(valid_indices) < cutoff) {
    warning(paste("Only", length(valid_indices), "points found within max_dist. Minimum required:", cutoff))
    return(list(
            prediction = NA,
            n_points_used = length(valid_indices),
            mean_distance = ifelse(length(valid_indices) > 0, mean(distances[valid_indices]), NA),
            weights = NA,
            distances = distances
    ))
  }

  # Subset to valid points
  filtered_distances <- distances[valid_indices]
  filtered_values <- values[valid_indices]

  # Calculate weights (inverse distance)
  weights <- 1 / (filtered_distances ^ power)

  # Handle case where prediction point coincides with a data point
  if (any(filtered_distances == 0)) {
    exact_match <- which(filtered_distances == 0)[1]
    prediction <- filtered_values[exact_match]
    weights <- numeric(length(filtered_distances))
    weights[exact_match] <- 1
  } else {
    # Normalize weights
    weights <- weights / sum(weights)

    # Calculate weighted average
    prediction <- sum(weights * filtered_values)
  }

  return(list(
          prediction = prediction,
          n_points_used = length(valid_indices),
          mean_distance = mean(filtered_distances),
          weights = weights,
          distances = distances
  ))
}


custom_st_distance <- function(x1, y1, t1, x2, y2, t2,
                               time_scale = 1,
                               space_weight = 0.5,
                               time_weight = 0.5) {
  # Calculate spatial distance (Euclidean)
  spatial_dist <- sqrt((x2 - x1)^2 + (y2 - y1)^2)

  # Calculate temporal distance with scaling
  temporal_dist <- abs(t2 - t1) * time_scale

  # Combine using weighted approach
  combined_dist <- space_weight * spatial_dist + time_weight * temporal_dist

  return(combined_dist)
}
```

```{r}
pred_loc <- c(73, 27.5)
pred_time <- as.POSIXct("2023-10-01")

idw_d1 <- st_idw(full_filtered, pred_loc, pred_time, c('longitude', 'latitude'), 'data_time', 'data_value', dist_func = custom_st_distance)
glimpse(idw_d1)
```
Convincing enough.

```{r}
x_min <- min(dataset$longitude, na.rm = TRUE)
x_max <- max(dataset$longitude, na.rm = TRUE)
y_min <- min(dataset$latitude, na.rm = TRUE)
y_max <- max(dataset$latitude, na.rm = TRUE)

t_min <- min(dataset$data_time, na.rm = TRUE)
t_max <- max(dataset$data_time, na.rm = TRUE)

x_grid <- seq(x_min, x_max, length.out = 10)  # Increase/decrease for finer/coarser resolution
y_grid <- seq(y_min, y_max, length.out = 10)
t_grid <- c(as.POSIXct("2023-08-01"), as.POSIXct("2023-12-01"), as.POSIXct("2024-04-01"))


# Initialize results matrix
predictions <- array(NA, dim = c(length(x_grid), length(y_grid), length(t_grid)))

# Perform predictions for visualization
for (t_idx in 1:length(t_grid)) {
  for (x_idx in 1:length(x_grid)) {
    for (y_idx in 1:length(y_grid)) {
      result <- st_idw(
              dataset = dataset,
              pred_loc = c(x_grid[x_idx], y_grid[y_idx]),
              pred_time = t_grid[t_idx],
              coords_cols = c("longitude", "latitude"),
              time_col = "data_time",
              value_col = "data_value",
              power = 2,
              time_scale = 0.5,
              dist_func = custom_st_distance
      )
      predictions[x_idx, y_idx, t_idx] <- result$prediction
    }
  }
}
```

Here's the formula for space-time cube distance in LaTeX:

$$d_{\text{ST}}(\mathbf{p}, \mathbf{q}) = \sqrt{\left(\frac{x_p - x_q}{s_x}\right)^2 + \left(\frac{y_p - y_q}{s_y}\right)^2 + \left(\frac{t_p - t_q}{s_t}\right)^2}$$

Where:
- $\mathbf{p} = (x_p, y_p, t_p)$ and $\mathbf{q} = (x_q, y_q, t_q)$ are two points in the space-time cube
- $(x_p, y_p, t_p)$ and $(x_q, y_q, t_q)$ are the spatial-temporal coordinates
- $s_x$, $s_y$, and $s_t$ are scaling factors for each dimension
- $s_x$ and $s_y$ normalize spatial units (e.g., meters, kilometers)
- $s_t$ normalizes temporal units (e.g., seconds, days) to make them comparable with spatial units

The scaling factors are crucial in the space-time cube approach as they explicitly address the different units and ranges of the spatial and temporal dimensions. These factors are often derived from the characteristic scales of the phenomena being studied or from the extent of the dataset.