---
title: "Sanity Checks on the GWL Dataset"
output: pdf_document
---

The first course of action is to analyse the dimensions of `gwl_2023_24.feather`.
```{r}
library(dplyr)
library(feather)
library(ggplot2)
library(rmarkdown)

gwl_2023_24 <- read_feather('../data/gwl_2023_24.feather')
glimpse(gwl_2023_24)
```
There are **4,308,184** rows of data for one year! This does not align with our ex-ante knowledge that the order
of count of stations is in the _ten thousands_, and around _4 observations_ are taken over one year.
```{r}
observations_per_station <- gwl_2023_24 %>%
  count(latitude, longitude, sort = TRUE)
glimpse(observations_per_station)
```
```{r}
summary(observations_per_station$n)
```
One half of our prior knowledge has been confirmed-there are _18,849 unique lat-long pairs_, each referring to a
groundwater level measuring station.
All upto the 3rd quartile, every latitude-longitude pair has a single-digit observation count.
What happens beyond this quartile is open to question!
```{r}
q3 <- quantile(observations_per_station$n, 0.75)
stations_beyond_75_percentile <- observations_per_station %>%
        filter(n > q3)
glimpse(stations_beyond_75_percentile)
```
As a test to make things simpler for ourselves, we will consider only the first station (30.60668, 75.88028) and see what
is going on with the datapoints from this station.

```{r}
most_observed_station <- gwl_2023_24 %>%
        filter(latitude == 30.60668, longitude == 75.88028) %>%
        arrange(data_time)
glimpse(most_observed_station)
```
The `glimpse()` function is showing us at a glance that the Punjab Barnala Mehla Kalan Chananwal station has been
taking readings at a much higher frequency than was expected! For good measure, we will confirm if the time stamps
are unique throughout those 12,924 readings:
```{r}
timestamp_count <- most_observed_station %>%
        summarize(count = n_distinct(data_time)) %>%
        pull(count)
timestamp_count
```
A fraction of those datapoints have unique timestamps! Are these duplicates?
```{r}
readings_per_time <- most_observed_station %>%
        group_by(data_time) %>%
        summarize(n = n(),
                  n_distinct = n_distinct(data_value),
                  mean = mean(data_value))
glimpse(readings_per_time)
```
Well.
In an unexpected turn of events, _almost all the different readings collected at the same timestamp have different values!_
A pointless side quest would be to identify the timestamps, although such information does not really help with the analysis.
```{r}
unequal_timestamps <- readings_per_time %>%
        filter(n != n_distinct) %>%
        arrange(desc(n - n_distinct))
glimpse(unequal_timestamps)
```
Moving on.
Another simple (oh, how naïve) effort would be to find all the readings on a random timestamp:
```{r}
random_datetime <- most_observed_station %>%
        filter(data_time < as.POSIXct('2023-05-01 06:00:00'))
summary(random_datetime$data_value)
```
_That is surely a large range of observations for a measurement that had been allegedly taken at the same location
and time!_ This is a conclusion on the quality of data collected at the Punjab Barnala Mehla Kalan Chananwal station.
Such an exercise can be extended to other stations as well, which also report an abnormally high amount of data.

# Possible Ways Out
## Averaging over spatio-temporal locations
A simple, although very questionable method to use from here on would be to simply average over the multiple readings
taken at the same spatio-temporal location.
```{r}
gwl_2023_24_reduced <- gwl_2023_24 %>%
        group_by(latitude, longitude, data_time) %>%
        summarise(mean_gwl = mean(data_value))
glimpse(gwl_2023_24_reduced)
```
And yet, our valiant 4-line effort does little to reduce the dataset size by much!
Let's continue the sanity checks that we had done previously:
```{r}
observations_per_station_reduced <- gwl_2023_24_reduced %>%
        group_by(latitude, longitude) %>%
        summarise(n = n()) %>%
        arrange(desc(n))
summary(observations_per_station_reduced$n)
```
There exists a station which has made 7089 (averaged over each timestamp) observations over the year!
```{r}
glimpse(observations_per_station_reduced)
```
```{r}
station_info <- gwl_2023_24 %>%
        filter(latitude == 28.707800 & longitude == 77.2490) %>%
        head(1)
glimpse(station_info)
```
There's our culprit. How's their observations coming along?
```{r}
sonia_vihar <- gwl_2023_24_reduced %>%
        filter(latitude == 28.707800 & longitude == 77.2490)
glimpse(sonia_vihar)
summary(sonia_vihar$mean_gwl)
```
But of course, all the timestamps will be unique due to our previous `group_by` operations.
```{r}
length(unique(sonia_vihar$data_time))
```
Here's a visualisation of the mean groundwater level in metres at Sonia Vihar, for no reason.
```{r}
ggplot(sonia_vihar, mapping = aes(y = mean_gwl, x = data_time)) +
        geom_line()
```


Initially thought to be a pointless exercise, the data visualisation has brought us a peculiar graph, and brings us
to ask stranger questions: what happened to the readings in September?

**In any case, the time-stamps are at the heart of the true reason as to why the dataset is this massive.**

### We will conclude here briefly until better ideas come up to manage this dataset.
- Aggregate data on time-stamps?