---
title: "Inverse Distance Weighing on the Filtered GWL Dataset"
output: pdf_document
---

For more information on how we obtained the balanced panel, kindly refer to `exploratory/filtered_panel.rmd`.
```{r}
# install.packages(c('dplyr', 'readr', 'gstat', 'spacetime', 'sp', 'xts', 'fields', 'lubridate', 'ggplot2', 'glue'))

library(dplyr)
library(readr)
library(gstat)
library(spacetime)
library(sp)
library(xts)
library(lubridate)
library(fields)
library(ggplot2)
library(glue)

gwl_panel <- read_csv("data/year_panel.csv")
gwl_panel <- gwl_panel %>%
  rename(x = longitude, y = latitude, time = data_time) %>%
  mutate(month_ = month(floor_date(time, unit = "month")))
```

We will perform **inverse distance weighing** using different distance functions and parameters.
First and foremost, we will be using Euclidean distance, using the code in _(Wikle et al., 2015)_
```{r}
# Euclidean Distance
euclidean <- function (theta) {
  all_results <- list()

  for (i in seq_along(alpha_values)) {
    alpha <- alpha_values[i]
    pred_grid <- expand.grid(x = seq(min(gwl_panel$x), max(gwl_panel$x), length = 50),
                             y = seq(min(gwl_panel$y), max(gwl_panel$y), length = 50),
                             month_ = 10)
    pred_obs_dist_mat <- rdist(select(pred_grid, x, y, month_),
                               select(gwl_panel, x, y, month_))

    # list of kernels to use
    Wt_IDW <- function(theta_, dist_mat) 1/dist_mat^theta_
    # Wt_Gauss <- function (theta_, dist_mat) exp(-dist_mat^2/theta_)

    Wtilde <- Wt_IDW(theta_ = alpha, dist_mat = pred_obs_dist_mat)
    Wtilde_rsums <- rowSums(Wtilde)
    W <- Wtilde/Wtilde_rsums
    z_pred_IDW <- as.numeric(W %*% gwl_panel$reading)

    all_results[[i]] <- pred_grid %>%
      mutate(prediction = z_pred_IDW,
             alpha = as.factor(alpha))
  }
  do.call(rbind, all_results)
}

euclidean_plot <- function (alpha_values) {
  pred_results <- euclidean(alpha_values)

  india <- ne_countries(scale = "medium", returnclass = "sf", country = "India")
  bbox <- st_bbox(c(xmin = min(pred_results$x),
                    xmax = max(pred_results$x),
                    ymin = min(pred_results$y),
                    ymax = max(pred_results$y)),
                  crs = st_crs(4326))
  bbox_poly <- st_as_sfc(bbox)

  # Crop India border to your prediction area
  india_cropped <- st_intersection(india, bbox_poly)
  delhi <- data.frame(
    x = 77.1025,
    y = 28.7041,
    label = "Delhi"
  )

  # pred_results <- pred_results %>% rename(theta = alpha)
  ggplot() +
    geom_tile(data = pred_results, aes(x = x, y = y, fill = prediction)) +
          geom_sf(data = india, fill = NA, color = "white", size = 2.5) +
    geom_point(data = delhi, aes(x = x, y = y),
               color = "white", size = 3, shape = 17) +
    geom_text(data = delhi, aes(x = x, y = y, label = label),
              hjust = -0.2, vjust = -0.5, color = "white", size = 5) +
    scale_fill_viridis_c(name = "Predicted \nLevel (m)") +
          coord_sf(xlim = c(min(pred_results$x), max(pred_results$x)),
                   ylim = c(min(pred_results$y), max(pred_results$y)), expand = FALSE)  +
    facet_wrap(~ alpha, labeller = label_both) +  # Create facets by alpha value
    theme_minimal() +
    labs(title = "Inverse Distance Weighing (Euclidean Distance)",
         x = "Longitude",
         y = "Latitude",
         fill = "Predicted Value") +
    theme(plot.title = element_text(size = 20, hjust = 0.5),
          axis.title.x = element_text(size = 18),
          axis.title.y = element_text(size = 18),
          axis.text.x = element_text(size = 10),
          axis.text.y = element_text(size = 10),
          legend.title = element_text(size = 18),
          legend.text = element_text(size = 10),
          strip.text = element_text(size = 18),
          plot.margin = margin(0, 0, 0, 0))
}

alpha_values <- c(1, 0.2, 0.01)
alpha_values <- c(2, 5, 10)
euclidean_plot(alpha_values)

ggsave("plots/idw_all_subzero.png", dpi = 150)

# function calls to previous versions without faceting
# alpha <- 1
# pred_results <- euclidean(alpha)
# euclidean_plot(pred_results, alpha)
```


It is time to use Gaussian kernel prediction as well, using the same Euclidean distance metric.
Oh wait, we could just use the above code with a small tweak!
```{r}
# Wt_Gauss <- function (theta_, dist_mat) exp(-dist_mat^2/theta_)
```