---
title: "Updated OLS Regression Using Categorical Variables"
output: pdf_document
---

For more information on how we obtained the balanced panel, kindly refer to `exploratory/filtered_panel.rmd`.
```{r}
# install.packages(c('readr', 'gstat', 'spacetime', 'sp', 'xts', 'fields', 'stargazer', 'ggplot2', 'maps'))

library(maps)
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(stargazer)

gwl_panel <- read_csv("data/year_panel.csv")
```


# Model Specification
We will be adding the following categorical/dummy variables to our existing analysis:
- `monsoon(s, t)` = 1 if `t` falls in the range June-September, 0 otherwise.
- `indo_gangetic(s, t)` = 1 if `s` falls in the Indo-Gangetic region, 0 otherwise
- `north_east(s, t)` = 1 if `s` falls in North-East India, 0 otherwise
- `deccan(s, t)` = 1 if `s` falls in the Deccan plateau, 0 otherwise
A visualisation will help us find the relevant locations, as well as find the coordinates that
we will use to assign dummy variable values.
```{r}
india_map <- map_data("world") %>%
  filter(region == "India")

ggplot() +
        labs(title = "Regions of Interest") +
  geom_polygon(data = india_map,
               aes(x = long, y = lat, group = group),
               color = "black", alpha = 0, size = 0.1) +
  geom_point(data = gwl_panel,
             aes(x = longitude, y = latitude), size = 0.5) +
  geom_rect(aes(xmin = 90, xmax = 96, ymin = 22, ymax = 28),
            fill = NA, color = "red", size = 0.5, alpha = 0) +
  geom_rect(aes(xmin = 73.50, xmax = 89.49, ymin = 23.5, ymax = 31.28),
            fill = NA, color = "red", size = 0.5, alpha = 0) +
  geom_rect(aes(xmin = 77, xmax = 81, ymin = 15, ymax = 20),
            fill = NA, color = "red", size = 0.5, alpha = 0) +
        annotate("text", x = 93, y = 28.5, label = "North-East Region",
                 color = "red", size = 4, fontface = "bold") +
        annotate("text", x = 81.495, y = 32, label = "Indo-Gangetic Region",
                 color = "red", size = 4, fontface = "bold") +
        annotate("text", x = 79, y = 20.5, label = "Deccan Region",
                 color = "red", size = 4, fontface = "bold") +
        theme(plot.title = element_text(size = 20, hjust = 0.5),
              axis.title.x = element_text(size = 18),
              axis.title.y = element_text(size = 18),
              axis.text.x = element_text(size = 10),
              axis.text.y = element_text(size = 10))

ggsave("plots/roi_all.png", dpi = 150)
```


We will now add the required variables for our regression analysis.
```{r}
ne <- data.frame(xmin = 90, xmax = 96, ymin = 22, ymax = 28)
indo_g <- data.frame(xmin = 73.50, xmax = 89.49, ymin = 23.5, ymax = 31.28)
deccan_plateau <- data.frame(xmin = 77, xmax = 81, ymin = 15, ymax = 20)

within_region <- function (x, y, region) {
  return ((x >= region$xmin & x <= region$xmax & y >= region$ymin & y <= region$ymax))
}

df <- gwl_panel %>%
  rename(x = longitude, y = latitude, time = data_time) %>%
  mutate(time = as.numeric(month(floor_date(time, unit = "month"))),
         long_time = x * time,
         lat_time = y * time,
         lat_long = x * y,
         monsoon = as.numeric(time <= 9 & time >= 6),
         indo_gangetic = as.numeric(within_region(x, y, indo_g)),
         north_east = as.numeric(within_region(x, y, ne)),
         deccan = as.numeric(within_region(x, y, deccan_plateau))
  )

  ols_model <- lm(reading ~ y + x + lat_time + long_time + lat_long +
        monsoon + indo_gangetic + north_east + deccan,
                data = df)

summary(ols_model)
# stargazer(ols_model, type = "text",
#           title = "OLS Regression Model Statistics", omit.stat = "all", report = "vcs*")
stargazer(ols_model, type = "latex",
          title = "OLS Regression Model Statistics", coef = NULL, p.auto = FALSE)

stargazer(ols_model,
          title = "Alternate OLS Regression Model Statistics",
          omit.table.layout = "n",
          type = "text")  # or "html" or "latex"
```


Time to visualise the residuals.
```{r}
df$residuals <- ols_model$residuals

df_chosen_times <- df %>%
        filter(month(time) %in% c(5, 9, 11, 3))

india_map <- map_data("world") %>%
        filter(region == "India")

ggplot() +
        # Add the map as a polygon layer
        geom_polygon(data = india_map,
                     aes(x = long, y = lat, group = group),
                     color = "black", alpha = 0, size = 0.1) +
        geom_point(data = df_chosen_times,
                   aes(x = x, y = y, color = residuals),
                   size = 0.15) +
        scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
        coord_fixed(1.3) +
        labs(title = "Spatial Plot of OLS Residuals over Months",
             x = "Longitude",
             y = "Latitude") +
        facet_wrap(~time, nrow = 1) +
        theme(plot.title = element_text(size = 20, hjust = 0.5),
              axis.title.x = element_text(size = 18),
              axis.title.y = element_text(size = 18),
              axis.text.x = element_text(size = 10),
              axis.text.y = element_text(size = 10),
              legend.title = element_text(size = 18),
              legend.text = element_text(size = 10),
              strip.text = element_text(size = 18),
              plot.margin = margin(0, 0, 0, 0))

ggsave('plots/ols_space_residuals.png', dpi = 150)
```
```{r}
sample_size <- 2
random_indices <- sample(seq_len(nrow(df)), size = sample_size)
random_latlong <- df[random_indices, c("y", "x")]

df_chosen_locations <- df %>%
        inner_join(random_latlong, by = c("y", "x"))

df_location_1 <- df_chosen_locations %>%
        filter(x == random_latlong$x[1])
df_location_2 <- df_chosen_locations %>%
        filter(x == random_latlong$x[2])

ggplot(df_location_1, aes(x = time, y = residuals)) +
        geom_line() +
        geom_hline(yintercept = 0, color = "red", linetype = "dashed") +  # Reference line at y=0
        labs(
                x = "Time",
                y = "Residuals",
                title = "Time Series Plot of OLS Residuals at Random Locations"
        ) + geom_line(data = df_location_2, mapping = aes(x = time, y = residuals)) +
        theme_bw() +
        theme(plot.title = element_text(size = 20, hjust = 0.5),
              axis.title.x = element_text(size = 18),
              axis.title.y = element_text(size = 18),
              axis.text.x = element_text(size = 15),
              axis.text.y = element_text(size = 15),
              legend.title = element_text(size = 18),
              legend.text = element_text(size = 10),
              strip.text = element_text(size = 18),
              plot.margin = margin(0, 0, 0, 0))

ggsave('plots/ols_time_residuals.png', dpi = 150)
```

Variogram megacope.
```{r}
# data version
coords <- df %>%
        group_by(x, y) %>%
        summarise()
coords <- cbind(coords$x, coords$y)
sp_obj <- SpatialPoints(coords)

df <- df %>% mutate(timestamp = gwl_panel$data_time)
dates <- df$timestamp
dates <- unique(dates)
# dates <- as.POSIXct(glue("2023-"))
time_obj <- xts(1:12, order.by=dates)

n_locations <- length(sp_obj)
n_times <- length(time_obj)

obs_data <- data.frame(
        reading = df$reading,
        residuals = df$residuals
)
stfdf_obj <- STFDF(sp = sp_obj, time = time_obj, data = obs_data)

summary(stfdf_obj)
```
```{r}
var_residuals <- variogram(object = residuals~1,
                           data = stfdf_obj,
                           width = 2,
                           cutoff = 5000,
                           tlags =  0:11)
var_og <- variogram(object = reading~1,
                    data = stfdf_obj,
                    width = 2,
                    cutoff = 5000,
                    tlags =  0:11)

# First, let's combine the variogram data with an identifier
var_residuals$type <- "Residuals"
var_og$type <- "Original"

# Combine the two variograms into one data frame
combined_var <- rbind(var_residuals, var_og)

ggplot(combined_var, aes(x = dist, y = gamma)) +
        geom_tile(aes(color = factor(timelag)), size = 2) +
        facet_wrap(~ type, scales = "free_y") +
        labs(
                title = "Comparison of Spatiotemporal Variograms",
                x = "Distance",
                y = "Semivariance",
                color = "Time Lag"
        ) +
        theme_bw() +
        scale_color_viridis_d() # Using viridis color palette for better distinction


# attempt 2
# First, determine the common range for both variograms
min_gamma <- min(min(var_og$gamma), min(var_residuals$gamma))
max_gamma <- max(max(var_og$gamma), max(var_residuals$gamma))

# Set up the plotting grid
par(mfrow = c(1, 2))

# Plot original data variogram with fixed zlim (color scale limits)
plot(var_og,
     main = "Empirical Semivariogram (Full Data)",
     col.regions = viridis(100, option = "plasma", direction = -1),
     zlim = c(min_gamma, max_gamma),  # Set fixed range for color scale
     as.table = TRUE,
     xlab = "Distance (km)",
     ylab = "time lag (days)")

# Plot residuals variogram with the same zlim
plot(var_residuals,
     main = "Empirical Semivariogram (Residuals)",
     col.regions = viridis(100, option = "plasma", direction = -1),
     zlim = c(min_gamma, max_gamma),  # Same fixed range
     as.table = TRUE,
     xlab = "Distance (km)",
     ylab = "time lag (days)")

# Reset plotting parameters
par(mfrow = c(1, 1))
```

```{r}
proj4string(stfdf_obj) <- CRS("+proj=longlat +datum=WGS84 +no_defs")

# Transform to UTM (choose zone based on your area)
stfdf_proj <- spTransform(stfdf_obj, CRS("+init=epsg:32644"))  # UTM 44N


max_dist <- max(dist(coordinates(stfdf_obj)))
cutoff_dist <- max_dist * 0.5


var_residuals <- variogram(object = residuals~1,
                           data = stfdf_proj,
                           width = 5000,     # 5 km bins
                           cutoff = 200000,  # 200 km maximum
                           tlags = 0:11)
var_og <- variogram(object = reading~1,
                           data = stfdf_proj,
                           width = 5000,     # 5 km bins
                           cutoff = 200000,  # 200 km maximum
                           tlags = 0:11)
# Extract spatial variogram (time lag = 0) from both variograms
spatial_var_residuals <- var_residuals[var_residuals$timelag == 0, ]
spatial_var_og <- var_og[var_og$timelag == 0, ]

# Combined plot with both variograms
ggplot() +
        geom_point(data = spatial_var_og, aes(x = dist, y = gamma, color = "Original Data")) +
        geom_line(data = spatial_var_og, aes(x = dist, y = gamma, color = "Original Data")) +
        geom_point(data = spatial_var_residuals, aes(x = dist, y = gamma, color = "Residuals")) +
        geom_line(data = spatial_var_residuals, aes(x = dist, y = gamma, color = "Residuals")) +
        scale_color_manual(values = c("Original Data" = "blue", "Residuals" = "red")) +
        scale_x_continuous(labels = function(x) x/1000) +  # Convert to km on axis
        labs(x = "Distance (km)", y = "Semivariance",
             title = "Spatial Variograms",
             color = "Data Type") +
        theme_bw() +
        theme(plot.title = element_text(size = 20, hjust = 0.5),
              axis.title.x = element_text(size = 18),
              axis.title.y = element_text(size = 18),
              axis.text.x = element_text(size = 15),
              axis.text.y = element_text(size = 15),
              legend.title = element_text(size = 18),
              legend.text = element_text(size = 12),
              strip.text = element_text(size = 18),
              plot.margin = margin(0, 0, 0, 0))

ggsave('plots/sp_variogram.png', dpi = 150)
```

```{r}
# Extract temporal variogram (typically using the smallest distance bin)
# Find the smallest distance bin
min_dist <- min(var_residuals$dist, na.rm = TRUE)

# Extract temporal variogram for residuals and original data
temporal_var_residuals <- var_residuals[var_residuals$dist == min_dist, ]
temporal_var_og <- var_og[var_og$dist == min_dist, ]

# Plot the temporal variograms
ggplot() +
  geom_point(data = temporal_var_og, aes(x = timelag, y = gamma, color = "Original Data")) +
  geom_line(data = temporal_var_og, aes(x = timelag, y = gamma, color = "Original Data")) +
  geom_point(data = temporal_var_residuals, aes(x = timelag, y = gamma, color = "Residuals")) +
  geom_line(data = temporal_var_residuals, aes(x = timelag, y = gamma, color = "Residuals")) +
  scale_color_manual(values = c("Original Data" = "blue", "Residuals" = "red")) +
  labs(x = "Time Lag", y = "Semivariance",
       title = "Temporal Variograms",
       color = "Data Type") +
  theme_bw() +
  theme(plot.title = element_text(size = 20, hjust = 0.5),
        axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12),
        strip.text = element_text(size = 18),
        plot.margin = margin(0, 0, 0, 0))

ggsave('plots/t_variogram.png', dpi = 150)
```



```{r}
library(gstat)


# Fit variogram model to spatial components
model_var_og <- fit.variogram(spatial_var_og, vgm(psill = max(spatial_var_og$gamma, na.rm = TRUE)/2,
                                                  model = "Exp",
                                                  range = max(spatial_var_og$dist, na.rm = TRUE)/3))

model_var_residuals <- fit.variogram(spatial_var_residuals, vgm(psill = max(spatial_var_residuals$gamma)/2,
                                                                model = "Exp",
                                                                range = max(spatial_var_residuals$dist)/3))

# Extract parameters
og_nugget <- model_var_og$psill[1]
og_sill <- sum(model_var_og$psill)
og_range <- model_var_og$range[2]  # Effective range for exponential model

res_nugget <- model_var_residuals$psill[1]
res_sill <- sum(model_var_residuals$psill)
res_range <- model_var_residuals$range[2]

# Plot with annotations
ggplot() +
  geom_point(data = spatial_var_og, aes(x = dist, y = gamma, color = "Original Data")) +
  geom_line(data = spatial_var_og, aes(x = dist, y = gamma, color = "Original Data")) +
  geom_point(data = spatial_var_residuals, aes(x = dist, y = gamma, color = "Residuals")) +
  geom_line(data = spatial_var_residuals, aes(x = dist, y = gamma, color = "Residuals")) +
  scale_color_manual(values = c("Original Data" = "blue", "Residuals" = "red")) +

  # Add horizontal lines for sills
  geom_hline(yintercept = og_sill, linetype = "dashed", color = "blue") +
  geom_hline(yintercept = res_sill, linetype = "dashed", color = "red") +

  # Add horizontal lines for nuggets
  geom_hline(yintercept = og_nugget, linetype = "dotted", color = "blue") +
  geom_hline(yintercept = res_nugget, linetype = "dotted", color = "red") +

  # Add vertical lines for ranges
  geom_vline(xintercept = og_range, linetype = "dashed", color = "blue") +
  geom_vline(xintercept = res_range, linetype = "dashed", color = "red") +

  # Add annotations
  annotate("text", x = max(spatial_var_og$dist) * 0.8, y = og_sill,
           label = paste("Sill =", round(og_sill, 2)), color = "blue", hjust = 0) +
  annotate("text", x = max(spatial_var_og$dist) * 0.8, y = res_sill,
           label = paste("Sill =", round(res_sill, 2)), color = "red", hjust = 0) +

  annotate("text", x = max(spatial_var_og$dist) * 0.8, y = og_nugget,
           label = paste("Nugget =", round(og_nugget, 2)), color = "blue", hjust = 0) +
  annotate("text", x = max(spatial_var_og$dist) * 0.8, y = res_nugget,
           label = paste("Nugget =", round(res_nugget, 2)), color = "red", hjust = 0) +

  annotate("text", x = og_range, y = max(spatial_var_og$gamma) * 0.9,
           label = paste("Range =", round(og_range, 2)), color = "blue", angle = 90, hjust = 1) +
  annotate("text", x = res_range, y = max(spatial_var_og$gamma) * 0.9,
           label = paste("Range =", round(res_range, 2)), color = "red", angle = 90, hjust = 1) +

  labs(x = "Distance", y = "Semivariance",
       title = "Spatial Variograms with Parameters",
       color = "Data Type") +
  theme_bw()
```


Statistical testing go!
```{r}
# Moran's I test
P <- list()
dates <- df$time
unique_dates <- unique(df$time)
for(i in unique_dates) {
  gwl_month <- filter(df, df$time == i)
  station.dists <- gwl_month %>%
    select(x, y) %>%
    dist() %>%
    as.matrix()

  station.dists.inv <- 1/station.dists
  diag(station.dists.inv) <- 0 #

  P[[i]] <- Moran.I(gwl_month$residuals, station.dists.inv) %>%
    do.call("cbind", .)
}

P <- do.call("rbind", P)
P <- as.data.frame(P)

glimpse(P)
summary(P)
sum(P$observed)
```


```{r}
# Calculate global Moran's I
# First create the distance matrix for all observations
station.dists <- df %>%
  select(x, y) %>%
  dist() %>%
  as.matrix()

# Create inverse distance weights matrix
station.dists.inv <- 1/station.dists
diag(station.dists.inv) <- 0  # Set diagonal to zero to avoid self-influence

# Calculate global Moran's I
global_moran <- Moran.I(df$residuals, station.dists.inv)

# Display the results
print(global_moran)
```

```{r}
# Durbin-Watson Test
nested_gwl <- group_by(df, x, y) %>% nest()
dwtest_one_station <- function(data)
  dwtest(residuals ~ 1, data = data)
map(nested_gwl$data, dwtest_one_station) %>% head()

gwl_DW <- nested_gwl %>%
  mutate(dw_test = map(data, dwtest_one_station)) %>%
  mutate(dw_statistic = map_dbl(dw_test, ~.$statistic),
         dw_pvalue = map_dbl(dw_test, ~.$p.value))


mean(gwl_DW$dw_statistic)
mean(gwl_DW$dw_pvalue)
mean(gwl_DW$dw_pvalue < 0.05/nrow(gwl_DW))
```
